---
title: "Functions"
author: "Austin Mercado"
date: "2023-03-21"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
convert_pounds_to_grams <- function(pounds = 1)
{grams <- 453.6 * pounds
return(grams)}
```

```{r}
convert_pounds_to_grams(pound = 10)
```

```{r}
get_mass_from_length_theropoda <- function(length, a, b)
{mass <- a * length ^ b
return(mass)}
```

#Error
get_mass_from_length_theropoda <- function(mass)
{mass <- a * length ^ b
return(mass)}

#Error
a = 214.44
b = 1.46
length <- 26
get_mass_from_length_theropoda()

```{r} 
get_mass_from_length_theropoda(length = 26, a =214.44, b = 1.46)
```

## Why do we use functions in programming?

- We already use functions a lot. Basically everything we run in R is a function. 
- Reduce the amount of code we write 
- No need to repeat code, so less chance of repeating errors or introducing new errors
by mistyping code. 
- Helps with reproducibility: no need to repeat code, allows others (and future self) to run code in a straightforward way, with accuracy.
- Make it easier to write modular code that you can reuse for other projects
- Makes it easier to run code and remember it.
- It allows for organization of analysis by putting code intended to run something together in the same function.

```{r}
convert_kg_to_pounds <- function(kg = 1) {pounds <- 2.205 * kg  
return(pounds)}
```

```{r}
convert_kg_to_pounds(kg = 50)
```

```{r}
convert_kg_to_pounds(kg = get_mass_from_length_theropoda(a = 10.95, b = 2.64, length = 12))
```

```{r}
length_in_pounds <- function(length, a, b){get_mass_from_length(length = length, a = a, b = b) %>% convert_kg_to_pounds()}
```

```{r}
length_in_pounds(length = 35, a = 13, b = 1.3)
```

Exercise 5: A function for the Portal data set

2. Combine the surveys and speices tables into a single data frame

There are several ways to create data frmaes within R. Depending on the data that we want as data frame, there are different functions that we can use to create a data frame object. 
- 'read.csv()': If you already have a file (either CSV or TSV) with data that we want to read into R as a data frame, 
- 'data.frame()': If we want to create a data frame from vectors
- '_join()' funtion: If we already have a some data frames in R that we want to combine.
- There are many functions that create a data frame as an output: 
    - Summarize(): It creates a new data frame summarizing information from another data frame.
    - but there are many other packages that also generate data frames with different types of data

3. Write a function that: 
Takes three arguments - a data frame (the combined table created before), a genus name, and a species name, 
Uses functions from the dplyr package to produce a data frame with two colums: year and cout, where count is the number of individuals (i.e., the number of rows) for the speices indicated by genus or species in that year, the functions return the resulting data frame. Note that this data frame is a time series by year. 

As opposed to data frames, there is one way to creat a function: the function 'function()':

'''{r}
name_of_the_function <- function(argument1, argument2, arguent3,...){The body of the function where the operation usng the vaulse is stored in the arguement}
'''

