---
title: "If-else conditions"
author: "Austin Mercado"
date: "2023-03-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Logical and conditional statements

These are pieces of code that reutnr the 'TRUE' or 'FALSE' values, that is, a logical value. 

THe common operators of logical statements are:
- equality '=='
- inequality '!='
- greater than '>'
- less than '<'
- greater or equal than '>='
- less or equal than '<='

The conditional statements allow to test several logical conditions at a time 
THe condition operators (or symbols) are
- AND '&' (inside dplyr function we can also represnt and using a ',')
- OR '|'

We also have logical functions that test if something is 'TRUE' or 'FALSE', for example:
- 'is.na()' is a function that exists if a value is an 'NA'
- This function is part of a whole family of functions, they all start with 'is.':
- is.vector()
- is.data.frame()
- is.factor()

For next class: how to get all functions from a family (method)

- which(): takes logical vectors, it will give you the numerical index (position) of all values that are TRUE

```{r}
which(letters == "r")
```

```{r}
w <- 10.2
x <- 1.3
y <- 2.8
z <- 17.5
colors <- c("red", "blue", "green")
masses <- c(45.2, 36.1, 27.8, 81.6, 42.4)
dna1 <- "attattaggaccaca"
dna2 <- "attattaggaacaca"
```

```{r}
w < 10
colors == "green"
x > y
masses > 40
2 * x + 0.2 == y
dna1 = dna2
dna1 != dna2
w > x | y > z
x*w == 13.2:13.5
masses == 30:50
```

```{r}
x * w < 13.5
x * w > 13.2
# 13.2 < x * w < 13.5. This is how we do it in paper
# But in R we have to compare things in pairs:
# For this we use the condtional statements
x * w < 13.5 & x * w > 13.2
```

## How to make simple choices with 'if()'

The general structure of an if statement:

```
if (condition is TRUE) {Run all lines of code in this block of code}
```

If the condition is not TRUE, then nothing happens.

```{r}
age_class = "sampling" 
if (age_class == "sampling") {
  y <- 10
} 
 y
```

```{r}
age_class = "seedling" 
if (age_class == "seedling") {
  y <- 10
} 
 y
```

## Handle more than 2 choices

In this case we are using the else if structure:
```
if (conditional) {first block code that is execites if condition 1 is met} else if (condition2) {second block code the executes if condition2 is met} else if (condition3) {more code} else {this will cover all the conditions that are not specified before}
```

You do not have to end up with and else block. 'Else if' are more intention the conditions. A simple else will run in all other cases no matter what. 

```{r}
if(age_class == "sapling") {
  y <- 10
}
if(age_class == "seedling") {
  y <- 5
}else if(age_class ==  "sampling") { y <- 10
} else {y <- 0}
y
```