---
title: "Joining Data Tables"
author: "Austin Mercado"
date: "2023-03-14"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

## R Markdown

```{r}
read.csv("../data-raw/plots.csv") -> plots 
read.csv("../data-raw/species.csv") -> species
read.csv("../data-raw/surveys.csv") -> surveys
```

```{r}
inner_join(surveys, species, by = "species_id")
```

```{r}
surveys |> inner_join(species, by = "species_id") -> joined_table
```

```{r}
head(species)
head(surveys)
head(joined_table)
```

```{r}
str(species)
str(surveys)
str(joined_table)
```

```{r}
surveys %>% inner_join(plots, by = "plot_id") %>% filter(plot_type == "Control")
```

```{r}
intersect(surveys$species_id, species$species_id)
```

```{r}
colnames(surveys)
```

```{r}
colnames(species)
```

```{r}
intersect(colnames(plots),
colnames(surveys))
```

```{r}
intersect(colnames(plots),
colnames(surveys))
```

```{r}
surveys %>% inner_join(plots, by = "plot_id") %>% filter(plot_type == "Rodent Exclosure")
```

## Other join functions
'left_join' retains all values from the first table, drops unmathching from second

'right_join' drops values from the first table and retaining all values from second

'full_join' keeps all values from both tables

## Joining multiple data tables

Can we use the '_join()' function on 3 or more tables at the same time?

No. It does tno recognize more than two tables at a time.

So we use a pipe and call the join function two or more times (as needed):

```{r}
inner_join(surveys, species, by = "species_id") %>% inner_join(plots, by = "plot_id") %>% str()
```

```{r}
inner_join(surveys, species, by = "species_id") %>% inner_join(plots, by = "plot_id") %>% filter(taxa == "Rodent") %>% select(year, genus, species, weight, plot_type) %>% filter(plot_type == "Control" | plot_type == "Long term Krat Exclosure") %>% filter(!is.na(weight))
```




