---
title: "trees-opentree"
author: "Austin Mercado"
date: "2023-04-20"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
install.packages("ape")
install.packages("rotl")
```
Insect - ~1 million named species
Mammals - ~5 thousands species 
Flowering plants - ~200K named species

```{r}
library("ape") # Analysis, Phylogenetics and Evolution R Package
library("rotl") #
library("nlme")
```

### Read a newick tree from the Open Tree of Life 

The following function can only read trees in newick format. Make sure your file is a newick tree. If you got your tree from the Phylesystem repo, your tree is going to be in json format

```{r}
tree <- read.tree(file = "../data-raw/subtree-node-ott247341-Canis-lupus.tre")
```

```{r, fig.height = 10, fig.width=5}
plot(tree, cex = 0.5)
mtext(text = "Canis Lupus OpenTree tree")
```

We can also use the url of a tree to read a tree into R. We will provie the eURL in the argument 'file ='.
In this case, we do not have to keep a copy of the file, we can just download it directly
The url is txt <- "http://ape-package.ird.fr/APER/APER2/primfive.tre":

```{r}
small_tree <- read.tree(file = "http://ape-package.ird.fr/APER/APER2/primfive.tre")
```

The R structure of a phylogenetic tree

### Getting a tree for your species from the Open Tree of Life

The Open Tree of Life has tools that allow to match the names of a group or species to their unified taxonomy so that we can look for information on their databases. 

If we are doing this in R, we are using the 'tnrs_match_names()' from the package 'rotl': tnrs stands for taxonmic name resolution service:

```{r}
tnrs <- tnrs_match_names(names = "amphibia")
class(tnrs)
tnrs
```

```{r}
tol_subtree(ott_id = tnrs$ott_id) -> amp_tree
```


The function above ran the TNRS service to match the common name that I provided to the Open Tree of Life Taxonomy (OTT) and get a unique numeric identifier called the OTT id.
For dolphins this is the number 'r tnrs_dolphins$ott_id'.
The next step is to use the OTT id in the function that gets a subtree from OpenTree:

```{r}
str(amp_tree)
class(amp_tree)
```

```{r}
plot.phylo(x = amp_tree,  cex = 200)
```
### Getting a tree for a set of taxa

Sometime we have very large groups and we just want to get a small tree representing some lineages inside the group. 

For example, we jsut want a tree of all orders in amphibians. We know the common names of these which are frogs, salamanders, and cecilians. 

We want to find this lineages in the Open Tree of Life; the steps to do this are:

1. Get the scientific names of our taxa 

```{r}
c("anura", "urodela", "gymophiona") -> amphibian_orders
```

2. Match the scientific names to the OTT to get a unique numeric identifier

```{r}
tnrs_match_names(names = amphibian_orders) -> amph_order_tnrs
amph_order_tnrs
```

We have two columns at the beginning, search string contains the names that we provided, and unique_name contains the most updated or official or most commonly used taxonomic names for the search taxon that we provided.

3. Next stp is to use the OTT ids that we got to extract the tree containing those lineages.
If we have only one name or lineage, we can use 'tol_subtree()'.
If we have more than one names, we will use 'tol_induced_subtree()'.

```{r}
tol_induced_subtree(ott_ids = amph_order_tnrs$ott_id) -> amp_order_tree
amp_order_tree
```

```{r}
plot.phylo(x = amp_order_tree)
```

What is the main difference between the information that we have in the column 'search_string' and 'unique_name':
One is the search string and the other is the scientific taxonomic name from the Open Tree of Life unified Taxonomy(OTT).

We can use the unique number identifier from OTT to get information from that taxon.

To get a tree we can use the function 'tol_subtree()'

```{r}
tol_subtree(ott_id = tnrs$ott_id)
```

