---
title: "UHURU Data Set Visualization"
author: "Austin Mercado"
date: "2023-02-23"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. A tip for working with Rmarkdown
The Working directory inside this Rmarkdown r chunk is the following: 
```{r eval=TRUE}
getwd()
```

Note: remember that working directories in a project and an R chunk are not always the same!

## 2. Describing the working data set

## 3. Reading the data set
```{r}
read.csv(file = "../data-raw/ACACIA_DREPANOLOBIUM_SURVEY.txt", sep = "\t")
```

Assign the data to a variable so we can work we can work with it
```{r}
ACACIA <- read.csv(file = "../data-raw/ACACIA_DREPANOLOBIUM_SURVEY.txt", sep = "\t")
```

## 4. Quality Control Check

```{r}
head(ACACIA)
str(ACACIA)
ACACIA
numbers <- 1:10
numbers
numbers[3:6]
numbers[c(1,5,6,3)]
```


```{r}
class(ACACIA$HEIGHT)
is.numeric(ACACIA$HEIGHT)
ACACIA$HEIGHT
as.numeric(ACACIA$HEIGHT)
```
We identified that HEIGHT should be numeric and is instead character:
```{r}
ACACIA$HEIGHT <- as.numeric(ACACIA$HEIGHT)
ACACIA$HEIGHT
```

Coercion in R computer language means that a value was forced to be type, in this case, the word "Dead" cannot be transformed to a number, so it is assigned the NA value
```{r}
ACACIA <- read.csv(file = "../data-raw/ACACIA_DREPANOLOBIUM_SURVEY.txt", sep = "\t", na.strings = "dead")
is.numeric(ACACIA$HEIGHT)
```

### 5. Accessing elements of a 'data.frame'

It is similar to what we do for vectors, but we have two dimensions 

```{r}
ACACIA[1,6]
str(ACACIA)
```

All the following are different ways of accessing a column or variable inside a data frame 
```{r}
ACACIA$SURVEY
ACACIA[,1]
ACACIA[,"SURVEY"]
```

### 6. Replacing elements in a data frame

```{r}
numbers[5] <- 100
numbers
```

## 7. Plotting data with 'ggplot2'
```{r}
plot(x = ACACIA$HEIGHT, y = ACACIA$AXIS1, main = "Acacia height vs Axis1")
?legend
legend(x = 1, y = 5, legend = "Acacia trees" )
```

With ggplot, we create layers
```{r}
library(ggplot2)

ggplot(data = ACACIA, mapping = aes(x = HEIGHT, y = AXIS1, color = TREATMENT)) + geom_point(size = 1 ) + labs(x = "Tree Height", y = "Tree Axis 1", title = "Acacia trees size relationships")
```

```{r}
colors()
View(ACACIA)
```

#####
Introductory section
#####
1. The goal of the experiment is to better understand how the interacting species corespond with and without one another in subsections of the areas the scientists have design to better understand the effect of one another(treatment).

2. (Images issue with pdf)

3. The variables that were measured were annual woody-plant censuses within the central 0.36 ha of each plot, annual and semi-annual monitoring of individually marked woody plants, small-mammal capture-mark-recapture sessions conducted eveyr other month in total-exclusion and open plots, daily rainfall monitoring throughout the course of the experiment and quarterly large-mammal dung surveys. 

4. (Graph maps aboved)

5. The influence resulted in varies of results that ranged from a wide range and some condensed within a interval of area; stunting it growth or decay. 

6. Those that remain in the open plots were stunted to a very low result, though on the graph it displays only one case of info, questioning on whether the others were invalid to report or were much lower to report.
