---
title: "Pipes"
author: "Austin Mercado"
date: "2023-03-07"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("dplyr")
library(magrittr)
surveys <- read.csv("../data-raw/surveys.csv")
```

```{r}
read.csv("../data-raw/surveys.csv") %>% # Read the data set
  select(year, month, day, species_id, weight) -> surveys1_pipe # select columns and assign to object

#name_object <- ode that we want to run
```


```{r}
surveys %>% select(year, species_id, weight) %>% mutate(weight_kg = weight/1000) %>% filter(!is.na(weight)) %>%
select(year, species_id, weight_kg) %>%
filter(species_id == "SH") -> surveys_final
str(surveys_final)
```

## Exercise 3

Reformat the following code in pipe mode:

```{r}
read.csv(file = "../data-raw/surveys.csv") %>%
  filter(species_id == "DS", !is.na(weight)) -> ds_data
str(ds_data)
ds_data %>% arrange(year) -> ds_data_by_year
str(ds_data_by_year)
ds_data_by_year %>% select(year, weight) -> ds_weight_by_year
str(ds_weight_by_year)
```

```{r}
read.csv(file = "../data-raw/surveys.csv") %>% filter(species_id == "DS", !is.na(weight)) %>% arrange(year) %>% select(year, weight) -> final_table
str(final_table)
head(final_table)
```

### Piping to an argument that is not the first one

Some functions do not take the data as the first arguement.
For example the 'lm()' function

```{r}
str(surveys)
lm(formula = weight ~ year, data = surveys)
```

```{r}
surveys %>% 
lm(formula = weight ~ year, data= .) # use an underscore if you are using this pipe |>
```

```{r}
read.csv("../data-raw/surveys.csv") %>% filter(species_id == "DS", !is.na(weight)) %>%  lm(weight ~ year, data= .) %>% summary()
```

### Grouping data or data aggregation

To get summary statistics for variables withing certain groups, we can group by our data by a certain value. For that we use the function 'group_by()'
```{r}
str(surveys)
group_by(.data = surveys, year)
grouped_surveys <- group_by(surveys, year)
str(grouped_surveys)
```

```{r}
group_by(surveys, year, sex)
```

### Get summary statistics of groups

The summary function create a new table that has the summary statistcs that we asked for per each group on the tibble.
```{r}
group_by(surveys, year, sex) %>% summarize(count = n())
```

```{r}
group_by(surveys, year) %>% summarize(mean_weight = mean(weight, na.rm = TRUE)) 
```

```{r}
read.csv("../data-raw/surveys.csv") %>% group_by(species_id) %>% summarize(count = n())
```

```{r}
read.csv("../data-raw/surveys.csv") %>% group_by(species_id, year) %>% summarize(count = n())
```

```{r}
read.csv("../data-raw/surveys.csv") %>% filter(species_id == "DO") %>% group_by(year) %>%  summarize(mean_mass = mean(weight, na.rm = TRUE)) 
```



